### 2022-02-24 structure.tex
smart clean

\usepackage[
nochapters,
beramono, 
eulermath
pdfspacing, 
dottedtoc 
]{classicthesis}

\usepackage{arsclassica}

\usepackage[T1]{fontenc}

\usepackage[utf8]{inputenc}

\usepackage{graphicx} 
\graphicspath{{Figures/}}

\usepackage{enumitem}

\usepackage{lipsum}

\usepackage{subfig}

\usepackage{amsmath,amssymb,amsthm}

\usepackage{varioref} 
\usepackage{siunitx}
\usepackage{makeidx}
\usepackage{minted}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{lscape}
\makeindex
\usepackage{glossaries}
\makeglossaries

\theoremstyle{definition} 
\newtheorem{definition}{Definition}

\theoremstyle{plain} 
\newtheorem{theorem}{Theorem}

\theoremstyle{remark}

\hypersetup{
%draft, % Uncomment to remove all links (useful for printing in black and white)
colorlinks=true, breaklinks=true, bookmarks=true,bookmarksnumbered,
urlcolor=webbrown, linkcolor=RoyalBlue, citecolor=webgreen, % Link colors
pdftitle={}, % PDF title
pdfauthor={\textcopyright}, % PDF Author
pdfsubject={}, % PDF Subject
pdfkeywords={}, % PDF Keywords
pdfcreator={pdfLaTeX}, % PDF Creator
pdfproducer={LaTeX with hyperref and ClassicThesis} % PDF producer
}